---
title: "Convention_appendices"
author: "Pauline"
date: '2022-05-06'
output: 
  html_document: 
    toc: TRUE
---
This file simply selects and exports all species present in conventions' appendices :  
* Barcelona_appendix : Convention for the Protection of the Mediterranean Sea Against Pollution (https://www.unep.org/unepmap/what-we-do). One assesment for elamsobranchs in 2007.
* BERN_appendix : Convention on the Conservation of European Wildlife and Natural Habitats (https://www.coe.int/en/web/bern-convention/home), indicator of status = appendix
* CITES : Convention on International Trade in Endangered Species of Wild Fauna and Flora (https://cites.org/eng), indicator of status = appendix
* CMS : Convention on the Conservation of Migratory Species of Wild Animals (https://www.cms.int/), indicator of status = appendix
* GROMS_migratory sp : database linked to CMS aka Bonn convention
* OSPAR : oslo-paris convention, own status by zone
* Shark.MoU : part of CMS dedicated to migratory sharks, indicator of status = appendix
* UNCLOS : United Nations Convention on the Law of the Sea, no species' assessments

```{r setup, include=FALSE}
#knitting options
knitr::opts_chunk$set(message=FALSE, warning=FALSE)

#Load packages
##general
library(tidyverse)
library(readxl)

##for the html output
library(knitr)
```

# Assesments from conventions
```{r conventions}
#elasmo_fr
elasmo_FR <- read.csv("data\\elasmo_uicn_france.csv")

#list of species names
species_names <- elasmo_FR$Species

#file with the corrected common names, used at the end of the doc for adding the common names
names_elasmo <- read_xlsx("data/Elasmobranches_présents_FR.xlsx") %>% rename(Species = Espèce, Name_FR = "Nom commun", Name_EN = "Nom anglais")

#oegelund nielsen's database
OegNiel_conventions <- read_excel("data/oegelund_nielsen_data.xlsx") %>% filter(Species %in% species_names) %>%
  select(Species, 11:20) %>%
  na_if("NA") %>% na_if(0) #for easier filtering after
```

## Conventions with appendices {#ConventionAppendices}
32 species are present in at least one appendix of Bern, Barcelona, CITES, Bonne
```{r appendices}
#only CITES has a listing year
OegNiel_appendices <- OegNiel_conventions %>%
  filter_at(vars(2:length(.)), any_vars(!is.na(.))) %>% #keep only species that are present in at least one appendix
  left_join(., names_elasmo[c("Species", "Name_FR", "Name_EN")], by = "Species") %>% #add common names
  relocate(c("Name_FR", "Name_EN"), .after = Species)

#export
write_excel_csv(OegNiel_appendices, file = "data/Species_cited_in_appendices.csv")
```

Remarks : I added by hand in the excel table whether they were part of the RFMOs' (ICCAT and GFCM, fishing organisations) listings.
