---
title: "06_Format_and_bind_all_tables"
author: "Pauline"
date: '2022-06-22'
output: html_document
---

# Purpose
Before creating the html version of the final matrix, you need to bind all the tables produced in the previous scripts to have your final df. To do so, you need to format them (same column names, same column types, etc. etc.). This is what this document does.

```{r setup, include = F}
library(tidyverse)
library(readxl)
library(writexl)

#select species (take 'commentaire' to have outdated names as well) for FC1 OSPAR
elasmo_FR <- read.csv("data\\elasmo_uicn_france.csv") %>% mutate(Genre = word(.$Species, 1))
species_names <- c(elasmo_FR$Species, elasmo_FR$Species) %>% unique() %>% na.exclude()
species_genre <- paste0(unique(elasmo_FR$Genre), " spp") %>% na.exclude()
species_all <- c(species_names, species_genre)

#file with the corrected common names, used at the end of the doc for adding the common names
names_elasmo <- read_xlsx("data/Elasmobranches_présents_FR.xlsx") %>% rename(Species = Espèce, Name_FR = "Nom commun", Name_EN = "Nom anglais")
```

# Load and format data
Logically, you load the documents produced at the very end of each script and stored in the data folder.
I mutate everything to character at the end to make sure the binding will work.

## UICN historical table
This one is ready for binding and it will be the model for the rest (for variable names etc.).
```{r data}
#IUCN historical table
Elasmobranchs_IUCN_historical_status <- read.csv("data\\Elasmobranchs_IUCN_historical_status.csv", encoding = "UTF-8", check.names = F) %>%
   mutate_all(as.character)

#last little correction to old statuses
Elasmobranchs_IUCN_historical_status[Elasmobranchs_IUCN_historical_status == "nt"] <- "NT"
Elasmobranchs_IUCN_historical_status[Elasmobranchs_IUCN_historical_status == "LR/nt"] <- "NT"
```

## Conventions
This one needs to have only one column for all appendices, separated by comas. The goal is to reduce information to bind it to the finale table.
```{r}
Appendices <- read.csv("data\\Species_cited_in_appendices.csv", check.names = F, sep = ",")

#Appendices table to long format
Appendices[Appendices == "1"] <- "" #turn 1s to empty character strings

Appendices_long <- Appendices %>% 
  select(-CITES_listing_year, -CITES_EU_listing, -CITES_party) %>% #we don't need these columns : for CITES keep only CITES_appendix
  rename(Bern = BERN_appendix, 
         Barcelona = Barcelona_appendix, 
         Washington = CITES_appendix, 
         "Bonn - CMS" = CMS_appendix,
         "Sharks MoU" = Shark.MoU) %>%
  mutate_all(as.character) %>%
  pivot_longer(cols = "Washington" : "Sharks MoU", names_to = "Conventions", values_to = "Détail", values_drop_na = T) %>%
  unite(Conventions:"Détail", col = Conventions, remove = TRUE, sep = " ", na.rm = T) %>%
  pivot_wider(values_from = "Conventions", names_from = "Conventions") %>% #group by wasn't working
  unite("Washington II" : "Washington I", col = Conventions, remove = T, sep = ", ", na.rm = T) %>%
   mutate_all(as.character)
```

## OSPAR
```{r}
#OSPAR Status
##POSH
Status_OSPAR <- read.csv("data\\Status_OSPAR.csv", check.names = F, sep = ";") #to have symbols

##FC1 --Got this one from the DCSMM team in the Museum, it hasn't been published yet
FC1_OSPAR <- readxl::read_xlsx("data\\FC1_integrated_outcomes.xlsx")

#Add OSPAR status for french waters (II to IV)
POSH <- Status_OSPAR %>% filter(Region %in% c("II", "III", "IV", "II*")) %>%
  mutate(Region = case_when( #rename regions
    Region == "II" | Region == "II*" ~ "OSPAR II - GNS (POSH)",
    Region == "III" ~ "OSPAR III - CS (POSH)",
    Region == "IV" ~ "OSPAR IV - BBIC (POSH)"
  )) %>%
  mutate("2021" = Last_status) %>% #add a column with year as name as in elasmo table
  select(-Year) %>% #and drop the year column
  select(-Common_name) %>%
  mutate(Level = "Regional") %>%
  mutate_all(as.character) %>%
  left_join(., names_elasmo[c("Species", "Name_FR","Name_EN")])
  

#FC1
FC1 <- FC1_OSPAR %>% filter(species %in% species_all) %>% filter(variable == "Long") %>% #keep only long term trend
  mutate(msfd.region = case_when(
    msfd.region == "BBIC" ~ "OSPAR IV - BBIC (FC1)",
    msfd.region == "GNS" ~ "OSPAR II - GNS (FC1)",
    msfd.region == "CS" ~ "OSPAR III - CS (FC1)"
  )) %>%
  mutate("2022" = bi.res) %>%
  rename("Species" = species, "Region" = msfd.region, "Last_status" = bi.res) %>% 
  select(Species, Region, Last_status, "2022") %>%
  mutate(Level = "Regional", Last_evaluation = "2022") %>%
  mutate(Last_status = str_to_title(Last_status)) %>%
  left_join(., names_elasmo[c("Species", "Name_FR", "Name_EN")]) %>%
   mutate_all(as.character)
```

## ICES
```{r}
#stock categories
ICES_stock_cat <- read_xlsx("data\\ICES_stock_cat.xlsx") %>% left_join(., names_elasmo[c("Species", "Name_FR", "Name_EN")]) %>% relocate(c("Name_FR", "Name_EN"), .after = Species)
write_xlsx(ICES_stock_cat, "data\\ICES_stock_cat.xlsx")

#Ices advice
ICES_advice <- read_xlsx("data\\ICES_advice_history_updated.xlsx") #drop 1st id col

#need to have this somewhere 
ICES_qualitative <- ICES_advice %>% filter(is.na(Stock_size_qualitative) == F) %>% select(Species, Stock_size_qualitative, Stock, "Eco Region", Year) %>%
  pivot_wider(names_from = Year, values_from = Stock_size_qualitative) %>%
  relocate(Stock, .before = Species) 

#export and add regions manually from the ecoregion text and the stock's name (sorry, didn't find a better way)
#writexl::write_xlsx(ICES_qualitative, "data\\ICES_qualitative_stock_eval.xlsx")

#import again
ICES_qualitative <- read_xlsx("data\\ICES_qualitative_stock_eval.xlsx") %>%
  select(-"Eco Region") %>%
  mutate(Region = case_when(
    Region == "NE Atlantic" ~ "NE Atlantic (ICES)",
    Region == "GNS" ~ "GNS (ICES)",
    Region == "CS (North)" ~ "Northern CS (ICES)",
    Region == "CS (South)" ~ "Southern CS (ICES)",
    Region == "BBIC (North)" ~ "Northern BBIC (ICES)",
    Region == "BBIC (South)" ~ "Southern BBIC (ICES)",
    Region == "CS" ~ "CS (ICES)"
  ),
  Level = "Regional",
  Last_status = ifelse(is.na(`2019`), `2018`, `2019`)) %>% 
  mutate(Last_status = str_to_title(Last_status)) %>%
  select(-Stock) %>%
 left_join(.,names_elasmo[c("Species", "Name_FR", "Name_EN")]) %>% relocate(c("Name_FR", "Name_EN"), .after = "Species") %>%
   mutate_all(as.character)

#finally replace inc and dec by full words
ICES_qualitative[ICES_qualitative == "Inc"] <- "Increasing"
ICES_qualitative[ICES_qualitative == "Dec"] <- "Declining"
ICES_qualitative[ICES_qualitative == "inc"] <- "Increasing"
ICES_qualitative[ICES_qualitative == "dec"] <- "Declining"
ICES_qualitative[ICES_qualitative == "stable"] <- "Stable"
ICES_qualitative[ICES_qualitative == "very low"] <- "Very low"
ICES_qualitative[ICES_qualitative == "depleted"] <- "Depleted"
```

## RFMOs
```{r}
#add ices and iccat
ICCAT_GFCM <- read.csv("data\\Status_ICCAT_GFCM.csv", sep = ";") %>%
  mutate(Last_evaluation = Year) %>%
  pivot_wider(values_from = Last_status, names_from = Year) %>%
  mutate(Region = case_when(
    Region == "NE Atlantic" ~ "NE Atlantic (ICCAT)",
    Region == "Ligurian and Tyrrhenian Seas" ~ "Ligurian and Tyrrhenian Seas (GFCM)" 
  ),
  Level = "Regional",
  Last_status = coalesce(`2019`, `2015`, `2009`, `2010`)) %>%
  select(-variable) %>%
  left_join(.,names_elasmo[c("Species", "Name_FR", "Name_EN")]) %>% relocate(c("Name_FR", "Name_EN"), .after = "Species") %>%
  mutate_all(as.character)
```

# Bind all together
```{r}
#do it step by step
Elasmobranchs_final_table <- full_join(Elasmobranchs_IUCN_historical_status, Appendices_long) %>% 
  relocate("Conventions", .after = "Name_EN") %>%
  full_join(., FC1) %>%
  full_join(., POSH) %>%
  full_join(., ICES_qualitative) %>%
   full_join(., ICCAT_GFCM) %>%
  drop_na(Last_status) %>% unique() %>% #make sure you have only valid rows
  arrange(Species, Level, Region, Last_status) %>%
  relocate(`2022`, .after = `2021`) %>% relocate(`2009`, .after = `2008`) #check that the year columns are in the right order !

#export the final table
write_xlsx(Elasmobranchs_final_table, "data\\Elasmobranchs_final_table.xlsx")
```
